SubsidyExplorer - Step 2: Create Vessel List

### 2020-08-06 scripts/02_vessel_list.Rmd
vessel list done

# Path to cleaned capture production data file
catch_dat_path <- here::here("data", "edited", "fao_2019_capture_production_tidy.csv")
catch_year <- 2017

# Path to cleaned landed value data file
landed_value_dat_path <- here::here("data", "edited", "estimated_landed_value_tidy.csv")
landed_value_year <- 2017

@@

# Do you wnat to make figures? 
make_figures <- FALSE
```

@@

```{r}
# Get table of all flag states in the data to check for weird flag codes
gfw_flags <- effort_good_fishing_vessels_eez %>%
  distinct(flag_iso3) %>%
  mutate(flag_name = countrycode(flag_iso3, "iso3c", "country.name"))

@@

catch_revenue_flag <- catch_revenue_flag_fao %>%
  group_by(iso3) %>%
  summarize(capture_production = sum(capture_production, na.rm = T),
            landed_value = sum(landed_value, na.rm = T))

@@

catch_no_matching_effort <- catch_revenue_flag %>%
  anti_join(effort_flag, by = c("iso3" = "flag_iso3"))

# There are 2 flag states for which we don't have reported catch data at all: CXR (Christmas Island), LIE (Liechtenstein) and NA (no flag). We're going to try assigning them based on the EEZ in which they spend the most time fishing instead. 
vessels_no_catches <- vessel_list %>%
  dplyr::filter(flag_iso3 %in% effort_no_matching_catches$flag_iso3) %>% # 1094 entries 
  group_by(ssvid, eez_id, eez_territory1_iso3, eez_territory2_iso3, eez_territory3_iso3) %>%
  summarize(fishing_KWh_eez_fao_ter = sum(fishing_KWh_eez_fao_ter, na.rm = T)) %>%
  group_by(ssvid) %>%
  mutate(rank = dense_rank(desc(fishing_KWh_eez_fao_ter)),
         max_rank_no_hs = min(rank[!is.na(eez_territory1_iso3)])) %>%
  ungroup()

@@

Now they all do and we can assign catches.

@@

# effort_flag
effort_no_matching_catches <- effort_flag %>%
  anti_join(catch_revenue_flag, by = c("flag_iso3" = "iso3"))
```

@@

# Matches
flag_region_matches <- effort_catch_region_matching %>%
  dplyr::filter(is_flag_region_match)

@@

### -------------------
comprehensive_catch_rates <- effort_catch_flag_region_matching %>%
  left_join(rates_flag_region_matches, by = c("flag_iso3", "fao_region")) %>%
  left_join(rates_flag_matches, by = c("flag_iso3")) %>%
  mutate(catch_per_fishing_KWh = case_when(!is.na(flag_region_catch_per_fishing_KWh) ~ flag_region_catch_per_fishing_KWh,
                                           TRUE ~ flag_catch_per_fishing_KWh),
         revenue_per_fishing_KWh = case_when(!is.na(flag_region_revenue_per_fishing_KWh) ~ flag_region_revenue_per_fishing_KWh,
                                           TRUE ~ flag_revenue_per_fishing_KWh)) %>%
  dplyr::select(flag_iso3, fao_region, catch_per_fishing_KWh, revenue_per_fishing_KWh)
```

@@

The process for allocating subsidies is slightly more straightforard as we only have data by flag

@@

# Summarize total subsidies by type and flag
subsidies_flag <- subsidies_dat %>%
  group_by(iso3, type) %>%
  summarize(value = sum(value, na.rm = T)) %>%
  spread(type, value)

@@

vessel_list <- vessel_list %>%
  mutate(catch = fishing_KWh_eez_fao_ter*catch_per_fishing_KWh,
         revenue = fishing_KWh_eez_fao_ter*revenue_per_fishing_KWh) %>%
  mutate_at(paste0(subsidy_types, "_per_fishing_KWh"), list(subs = ~.*fishing_KWh_eez_fao_ter)) %>%
  rename_at(paste0(subsidy_types, "_per_fishing_KWh_subs"), rename_me) %>%
  dplyr::select_at(vars(-contains("_per_fishing_KWh"))) %>%
  mutate(good_subs = rowSums(select(., one_of(paste0(good_subsidy_types, "_subs")))),
         bad_subs = rowSums(select(., one_of(paste0(bad_subsidy_types, "_subs")))),
         ugly_subs = rowSums(select(., one_of(paste0(ugly_subsidy_types, "_subs")))))

write_csv(vessel_list, paste0(results_dir, "vessel_list_2018_raw.csv"))
```

@@

```{r}
if(make_figures){
# Summary by gear type
vessels_by_gear <- vessel_list %>%
  group_by(vessel_class) %>%
  summarize(count = n_distinct(ssvid),
            flag_states = n_distinct(flag),
            fao_regions = n_distinct(fao_region),
            length_avg = mean(length_m, na.rm = T),
            tonnage_avg = mean(tonnage_gt, na.rm = T),
            engine_power_avg = mean(engine_power_kw, na.rm = T),
            fishing_hours = sum(fishing_hours_eez_fao, na.rm = T))

@@

# Total number of vessels
ggplot(vessels_by_gear)+
  aes(x = "", y = count, fill = vessel_class)+
  geom_bar(stat = "identity", width = 1, color = "grey4", size = 0.2)+
  coord_polar("y")+
  #scale_fill_tableau(name = "Flag State", palette = "Tableau 20")+
  pie_theme+
  labs(title = paste0("Number of vessels by vessel class | Total vessels: ", format(sum(vessels_by_gear$count), big.mark = ",")))

>>>>

# Path to cleaned ssf subsidy data file (2018)
ssf_subsidy_prop_dat_path <- here::here("data", "edited", "schuhbauer_ssf_subsidies_proportions.csv")

# Path to cleaned capture production data file
catch_dat_path <- here::here("data", "edited", "fao_2020_capture_production_tidy.csv")
catch_year <- 2018

# Path to cleaned landed value data file
landed_value_dat_path <- here::here("data", "edited", "estimated_landed_value_tidy.csv")
landed_value_year <- 2018

@@

# Do you wnat to make figures? 
make_figures <- FALSE

# Temporary solution to bigrquery error (August 5, 2020) - see https://github.com/r-dbi/bigrquery/issues/395 for details
options(scipen = 20)
```

@@

```{r}
# Get table of all flag states in the data to check for weird flag codes
gfw_flags <- effort_good_fishing_vessels_eez %>%
  distinct(flag_iso3) %>%
  mutate(flag_name = countrycode(flag_iso3, "iso3c", "country.name")) %>%
  arrange(flag_iso3)

@@

catch_revenue_flag <- catch_revenue_flag_fao %>%
  group_by(iso3) %>%
  summarize(capture_production = sum(capture_production, na.rm = T),
            landed_value = sum(landed_value, na.rm = T))
```

@@

# There are some vessels for which we don't have flags. We're going to try assigning them based on the EEZ in which they spend the most time fishing instead. 
vessels_no_catches <- vessel_list %>%
  dplyr::filter(flag_iso3 %in% effort_no_matching_catches$flag_iso3) %>% # 1094 entries 
  group_by(ssvid, eez_id, eez_territory1_iso3, eez_territory2_iso3, eez_territory3_iso3) %>%
  summarize(fishing_KWh_eez_fao_ter = sum(fishing_KWh_eez_fao_ter, na.rm = T)) %>%
  group_by(ssvid) %>%
  mutate(rank = dense_rank(desc(fishing_KWh_eez_fao_ter)),
         max_rank_no_hs = min(rank[!is.na(eez_territory1_iso3)])) %>%
  ungroup()

@@

Now we can assign catches.

@@

# Verify again 
effort_no_matching_catches <- effort_flag %>%
  anti_join(catch_revenue_flag, by = c("flag_iso3" = "iso3"))
```

@@

# Flag state-fao region pairings for which we have matching catch and effort data
flag_region_matches <- effort_catch_flag_region_matching %>%
  dplyr::filter(is_flag_region_match)

@@

### -------------------
# Apply catch/revenue rates by flag state-fao region pairings where available, otherwise use flag state rates
comprehensive_catch_rates <- effort_catch_flag_region_matching %>%
  left_join(rates_flag_region_matches, by = c("flag_iso3", "fao_region")) %>%
  left_join(rates_flag_matches, by = c("flag_iso3")) %>%
  mutate(catch_per_fishing_KWh = case_when(!is.na(flag_region_catch_per_fishing_KWh) ~ flag_region_catch_per_fishing_KWh,
                                           TRUE ~ flag_catch_per_fishing_KWh),
         revenue_per_fishing_KWh = case_when(!is.na(flag_region_revenue_per_fishing_KWh) ~ flag_region_revenue_per_fishing_KWh,
                                           TRUE ~ flag_revenue_per_fishing_KWh)) %>%
  dplyr::select(flag_iso3, fao_region, catch_per_fishing_KWh, revenue_per_fishing_KWh)
```

@@

The process for allocating subsidies is slightly more straightforard as we only have data by flag. However we

@@

# Load proportions going to SSF, and remove
subsidies_dat_lsf <- read_csv(ssf_subsidy_prop_dat_path) %>%
  dplyr::select(-source) %>%
  left_join(subsidies_dat, by = c("iso3", "category", "category_name", "type", "type_name")) %>%
  mutate(value_lsf = value*(1-prop_ssf))

# Summarize total subsidies by type and flag
subsidies_flag <- subsidies_dat_lsf %>%
  group_by(iso3, type) %>%
  summarize(value = sum(value_lsf, na.rm = T)) %>%
  spread(type, value)

@@

vessel_list_out <- vessel_list %>%
  mutate(catch = fishing_KWh_eez_fao_ter*catch_per_fishing_KWh,
         revenue = fishing_KWh_eez_fao_ter*revenue_per_fishing_KWh) %>%
  mutate_at(paste0(subsidy_types, "_per_fishing_KWh"), list(subs = ~.*fishing_KWh_eez_fao_ter)) %>%
  rename_at(paste0(subsidy_types, "_per_fishing_KWh_subs"), rename_me) %>%
  dplyr::select_at(vars(-contains("_per_fishing_KWh"))) %>%
  mutate(good_subs = rowSums(select(., one_of(paste0(good_subsidy_types, "_subs")))),
         bad_subs = rowSums(select(., one_of(paste0(bad_subsidy_types, "_subs")))),
         ugly_subs = rowSums(select(., one_of(paste0(ugly_subsidy_types, "_subs")))))

write_csv(vessel_list_out, paste0(results_dir, "vessel_list_2018_raw.csv"))
```

@@

```{r}
if(make_figures){
# Summary by gear type
vessels_by_gear <- vessel_list_out %>%
  group_by(vessel_class) %>%
  summarize(count = n_distinct(ssvid),
            flag_states = n_distinct(flag_iso3),
            fao_regions = n_distinct(fao_region),
            length_avg = mean(length_m, na.rm = T),
            tonnage_avg = mean(tonnage_gt, na.rm = T),
            engine_power_avg = mean(engine_power_kw, na.rm = T),
            fishing_hours = sum(fishing_hours_eez_fao_ter, na.rm = T),
            fishing_KWh = sum(fishing_KWh_eez_fao_ter, na.rm = T))

@@

# Total number of vessels
n_gear_type_plot <- vessels_by_gear %>%
  mutate(vessel_class = fct_re)
ggplot(vessels_by_gear)+
  aes(x = vessel_class, y = count, fill = vessel_class)+
  geom_bar(stat = "identity", width = 1, color = "grey4", size = 0.2)+
  #coord_polar("y")+
  #scale_fill_tableau(name = "Flag State", palette = "Tableau 20")+
  normal_theme+
  labs(title = paste0("Number of vessels by class | Total vessels: ", format(sum(vessels_by_gear$count), big.mark = ",")))+
  labs(x = "Gear Type", y = "Number of Vessels")+
  guides(fill = "none")+
  coord_flip()+
  scale_y_continuous(expand = c(0,0))
